---
title: "Creating plots from msleep"
author: "Cait Kaulens"
date: 
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
    options(tinytex.verbose = TRUE)
```


```{r}
install.packages("tidyverse")
install.packages("magick")
install.packages("tinytex")
install.packages("patchwork")
library(tidyverse)
library(dplyr)
library(ggplot2)
library(patchwork)
library(magick)
```


```{r, include=FALSE}
options(tinytex.verbose = TRUE)
```
\begin{center}
\textbf{\LARGE I am using the Mammals Sleep Dataset from Tidyverse to create some plots.}
\end{center}

\bigbreak

## Creating our tables for the following graphs. . .  

First, for anyone not familiar with this dataset we will take a brief look at the original information.  
  
  

```{r Original Data, echo=FALSE}
glimpse(msleep)
```
  
\bigbreak 
  
\textbf{Then,} I will start by selecting just a few of the columns so I can focus on the size of the animal compared to their diet and sleep patterns. I am also changing the weight measure from kilograms to pounds. 

```{r echo=TRUE}
sleep <- msleep %>% 
  select(name,vore,awake,sleep_total,bodywt)%>% 
  drop_na(vore) %>% 
  mutate(wt_lbs = (bodywt * 2.2)) %>% 
  arrange(-wt_lbs)
```
 

## After creating the sleep table, 
### we will modify it a bit to show averages of body weight and sleep totals.

```{r echo=TRUE}
gsleep <- sleep %>% 
  drop_na(vore) %>% 
  group_by(vore) %>% 
  summarise(Lower = min(sleep_total),
            Average = mean(sleep_total),
            Upper = max(sleep_total),
            Weight = mean(wt_lbs))%>% 
  arrange(Average)
```

  
\bigbreak 

\begin{center}
\textbf{\LARGE Let's get to the visualizations!}
\end{center}

\bigbreak

\begin{center}
The first plot will show Sleep Total and Body Weight by each kind of diet. 
\end{center}  

\bigbreak
 

```{r echo=FALSE, message=FALSE, results='hide', fig.align='center'}
  ggplot(sleep,aes(sleep_total,wt_lbs, color = vore))+
  geom_point(shape = 8)+
  scale_color_manual(values = c("#4F9FCE", "#A5165D","#A5F37D","#D25512"))+
  facet_wrap(~vore, scales = 'free')+
  scale_x_continuous(breaks = seq(0,20, by = 5))+
  labs(title = "Sleep Totals & Body Weight by ~Vore",
       x= "Hours of Sleep",
       y= "Body Weight (lbs)",
       caption = "data from famous msleep"
       )+
  theme_bw()+
  theme(plot.title = element_text(face = 'bold',
        hjust = 0.5),
        plot.caption = element_text(face = 'italic'),
        strip.text = element_text(face = 'bold', 
                                  color = "#34495E"),
        strip.background = element_rect(fill ="#EAF2F8"),
        legend.position = "none")

```


\newpage
## The next plot 
Shows a single point for each our mammal types, representing the average weight of each along with the average hours they spend asleep.
 
\bigbreak

```{r echo=FALSE}
gsleep%>% 
  ggplot(aes(Average,Weight))+
  geom_point(size = 4,shape = 18, aes(color = vore,))+
  scale_x_continuous()+
  labs(title = "Averages",
       x= "Avg Sleep by Diet Type",
       y= "Avg Weight by Diet Type")+
  scale_color_manual(values = c("#4F9FCE", "#A5165D","#A5F37D","#D25512"))+
  theme_bw()+
  theme(plot.title = element_text(face = 'bold',
        hjust = 0.5),
        panel.grid.major = element_line('white', size = 0.5),
        panel.grid.minor = element_blank())
```



\newpage
# And then. . .
combining the information into one viz, we can see exactly where the average weight & sleep fall within the individual points.
 
\bigbreak

```{r echo=FALSE}
ggplot(sleep,aes(sleep_total,wt_lbs))+
  geom_jitter(size = 1.5, aes(color = vore))+
  geom_point(data=gsleep,aes(Average,Weight, color = vore),
             size = 3.5, shape = 13,
             )+
  facet_grid(vore~., scales = 'free')+
  labs(x = "Sleep",
       y= "Body Wt (lbs)")+
  scale_color_manual(values = c("#4F9FCE", "#A5165D","#A5F37D","#D25512"), 
                    name = c("Avg Wt & Sleep"),
                    labels = c("carnivore","herbivore",
                               "insectivore","omnivore"))+
  theme_bw()+
  theme(strip.text = element_text(face = 'bold'),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
```

\newpage

\begin{center}
\textbf{\Large Looking at the variance of sleep totals by each diet type.}
\end{center}  
  
\bigbreak

```{r include=FALSE}
MinMax <-gsleep 
```
  
  

## Let's have some fun with different types of plots! 
I will show the minimum and maximum hours by each diet with a violin plot combined with the individual points by mammal.  

\bigbreak

```{r echo=TRUE, fig.keep='all', cache=TRUE}
 ggplot(sleep,aes(vore,sleep_total,fill=vore))+
  geom_violin(trim=FALSE,)+
  geom_dotplot(binaxis='y', stackdir='center',
                 position=position_dodge(1))+
  labs(title = "Difference in Sleep Totals",
       caption = "data from famous msleep",
       x= "-Vore",
       y= "Hours of Sleep")+
  scale_fill_manual(values = c("#4F9FCE", "#A5165D","#A5F37D","#D25512"))+
  theme_bw()+
  theme(text = element_text('Courier'),
        plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(face = 'italic',
                                    color = "#581845"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = 'none')
```

\bigbreak

\begin{center}
\textbf{\Large Let's see the average hours asleep and average hours awake side-by-side using patchwork.}
\end{center}  

\bigbreak

Taking a moment to manipulate the data even further. Earlier, I did this step with the time spent asleep. Now I am doing the same for the time spent awake.  

```{r echo=TRUE}
gawake <- sleep %>% 
  drop_na(vore) %>% 
  group_by(vore) %>% 
  summarise(Lower = min(awake),
            Average = mean(awake),
            Upper = max(awake),
            Weight = mean(wt_lbs))%>% 
  arrange(Average)
```


```{r}
Awake <- gawake
```
  
  
### Now we can look at the comparison!  


```{r echo=TRUE}
plot1 <- ggplot(MinMax,aes(vore,Average))+
  geom_point(aes(color = vore),
             size = 4,
             shape =8 )+
  geom_errorbar(aes(ymin = Lower, ymax = Upper, color = vore))+
  labs(title = "Avg Hours Asleep",
       caption = "data from famous msleep",
       x= "-Vore",
       y= "Hours of Sleep")+
  scale_color_manual(values = c("#4F9FCE", "#A5165D","#A5F37D","#D25512"))+
  theme_bw()+
  theme(text = element_text('Courier'),
        plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(face = 'italic',
                                    color = "#581845"),
        legend.position = 'none')

plot2 <- ggplot(Awake,aes(vore,Average))+
  geom_point(aes(color = vore),
             size = 4,
             shape =13)+
  geom_errorbar(aes(ymin = Lower, ymax = Upper, color = vore))+
  labs(title = "Avg Hours Awake",
       caption = "data from famous msleep",
       x= "-Vore",
       y= "Hours Awake")+
  scale_color_manual(values = c("#4F9FCE", "#A5165D","#A5F37D","#D25512"))+
  theme_bw()+
  theme(text = element_text('Courier'),
        plot.title = element_text(hjust = 0.5),
        plot.caption = element_text(face = 'italic',
                                    color = "#581845"),
        legend.position = 'none')
```

\bigbreak

```{r Side By Side, echo=FALSE}
plot1+plot2
```

